Uncertainty propagation in deep neural networks for sparse coding

### 2018-03-10 paper/BayesianLISTA.tex
trying to interpret results

\section{\uppercase{Neural networks for sparse coding}}
\label{sec:nn_sc}
The sparse coding and sparse inverse problems are the linear regression with the assumption of sparse weights
\begin{equation}
\label{eq:regression_problem}
\mathbf{y} = \mathbf{X}\boldsymbol\beta + \varepsilon
\end{equation}
where $\mathbf{y} \in \mathbb{R}^K$ is the observation, $\mathbf{X} \in \mathbb{R}^{K \times D}$ is the design matrix, $\boldsymbol\beta \in \mathbb{R}^D$ is the unknown vector of weight with many elements $\beta_d$ equal to zero. $\varepsilon \in \mathbb{R}^K$ is the independent Gaussian noise.

@@

\section{\uppercase{Experiments}}
\label{sec:experiments}
The proposed Bayesian \textsc{lista} is evaluated in the context of the sparse inverse problem, where the number of measurements $K$ is much smaller than the dimensionality of the vector $\boldsymbol\beta$.

We compare the proposed Bayesian \textsc{lista} with the classical \textsc{lista}~\citep{gregor2010learning} in terms of the predictive accuracy. As baselines we also use \textsc{ista} \citep{daubechies2004iterative} and Fast \textsc{ista} (\textsc{fista}) \citep{beck2009fast}. \textsc{fista} adds the momentum to \textsc{ista} and improves its convergence speed. We limit the number of iterations in these algorithms with $L$ that is the same as the number of layers in Bayesian \textsc{lista} and \textsc{lista}. To measure the performance of the algorithms the normalised mean square error(\textsc{nmse}) and F measure are used. \textsc{nmse} for a batch of data $\{\boldsymbol\beta^{(n)}\}_{n=1}^{N}$ and estimators $\{\widehat{\boldsymbol\beta}^{(n)}\}_{n=1}^{N}$ is computed as
\begin{equation}
\text{\textsc{nmse}} = \frac{1}{N}\sum\limits_{n=1}^N\sqrt{\frac{\sum\limits_{d=1}^D\left(\widehat{\beta}_{d}^{(n)} - \beta_d^{(n)}\right)^2}{\sum\limits_{d=1}^D\left(\beta_{d}^{(n)}\right)^2}}
\end{equation}
In sparse coding it is usually more important to obtain the correct locations of spikes (i.e zeros) and slabs (i.e. non-zeros) in the estimator than to minimise \textsc{nmse}. The problem is therefore viewed as a skewed two-class classification problem where the number of spikes is higher than the number of slabs. F measure is used to evaluate the accuracy of such problems. 
F measure is computed as the harmonic mean of precision and recall
\begin{equation}
\text{F measure} = 2\dfrac{\text{precision}\cdot\text{recall}}{\text{precision} + \text{recall}}
\end{equation} 
where precision is the fraction of estimated slab locations that are correct, recall is the fraction of true slab locations among all predicted slab locations.
We demonstrate the performance on small datasets to highlight that the proposed algorithm can infer accurate predictions when the dataset size is not sufficient for \textsc{lista} to learn anything.

@@

In the Figure \ref{fig:number_of_layers_synthetic} prediction performance for different number of layers $L$ is presented. The observation size is set as $K=50$.

In the Figure \ref{fig:unsersampling_synthetic} prediction performance for different observation sizes $K$ is presented. The number of layers is set as $L=4$.

@@

Figures \ref{fig:mnist_100} and \ref{fig:mnist_250} present quality on the validation set with dictionaries of size $100$ and $250$.

@@

%\begin{figure}[t]
%\includegraphics[width=\columnwidth]{loss}
%\caption{Validation \textsc{nmse}}
%\label{fig:validation}
%\end{figure}
\begin{figure*}[t]
\subfloat[$\beta$ posterior mean]{\includegraphics[width=0.66\columnwidth]{posterior_mean}}
\subfloat[$\beta$ posterior std]{\includegraphics[width=0.66\columnwidth]{posterior_std}}
\subfloat[$\beta$ posterior spike indicator]{\includegraphics[width=0.66\columnwidth]{posterior_spike_indicator}}
\caption{Posterior for the digit 7.}
\label{fig:posterior_distribution}
\end{figure*}
\begin{figure*}[t]
\subfloat[]{\includegraphics[width=0.66\columnwidth]{posterior_sample_0}}
\subfloat[]{\includegraphics[width=0.66\columnwidth]{posterior_sample_1}}
\subfloat[]{\includegraphics[width=0.66\columnwidth]{posterior_sample_2}}
\caption{Samples from the posterior for the digit 7.}
\label{fig:posterior_samples}
\end{figure*}

@@

The performance on the actively updated train data is compared with the performance on the randomly updated train data, when at every iteration the random element is moved from pool set to train set. Both algorithms of new data sampling were run with $20$ random seeds and their performance was averaged. Figure \ref{fig:active_learning_mnist} demonstrates their performance.
\begin{figure}[t]
\subfloat[\textsc{nmse}]{\includegraphics[width=\columnwidth]{graphics/active_mnist/nmse_validation}} \\
\subfloat[F measure]{\includegraphics[width=\columnwidth]{graphics/active_mnist/f_measure_validation}}
\caption{Active learning on \textsc{mnist}.}
\label{fig:active_learning_mnist}
\end{figure}

\section{\uppercase{Discussion}}
\label{sec:discussion}
As far as the authors aware, this is the first implementation of Bayesian deep sparse coding algorithm. Although there are works on Bayesian sparsity in context of neural networks \citep{he2017bayesian}, they are not the Bayesian neural networks in the same sense as Bayesian \textsc{lista} but rather the interpretation of the sparse Bayesian  learning algorithm as the long short-term memory network. We find not only correct predictions but also useful posterior estimates for the predictions that show how the model is confident in its decisions.

>>>>

\section{\uppercase{Neural networks for sparse coding}}
\label{sec:nn_sc}
The sparse coding problem is the linear regression with the assumption of sparse weights
\begin{equation}
\label{eq:regression_problem}
\mathbf{y} = \mathbf{X}\boldsymbol\beta + \varepsilon
\end{equation}
where $\mathbf{y} \in \mathbb{R}^K$ is the observation, $\mathbf{X} \in \mathbb{R}^{K \times D}$ is the design matrix, $\boldsymbol\beta \in \mathbb{R}^D$ is the unknown vector of weight with many elements $\beta_d$ equal to zero. $\varepsilon \in \mathbb{R}^K$ is the independent Gaussian noise.

@@

\section{\uppercase{Experiments}}
\label{sec:experiments}
The proposed Bayesian \textsc{lista} is evaluated in the context of the sparse coding problem with underdefined dictionary, where the number of measurements $K$ is much smaller than the dimensionality of the vector $\boldsymbol\beta$.

We compare the proposed Bayesian \textsc{lista} with the classical \textsc{lista}~\citep{gregor2010learning} in terms of the predictive accuracy. As baselines we also use \textsc{ista} \citep{daubechies2004iterative} and Fast \textsc{ista} (\textsc{fista}) \citep{beck2009fast}. \textsc{fista} adds the momentum to \textsc{ista} and improves its convergence speed. We limit the number of iterations in these algorithms with $L$ that is the same as the number of layers in Bayesian \textsc{lista} and \textsc{lista}. To measure the performance of the algorithms the normalised mean square error(\textsc{nmse}) and F measure are used. \textsc{nmse} for a batch of data $\{\boldsymbol\beta^{(n)}\}_{n=1}^{N}$ and estimators $\{\widehat{\boldsymbol\beta}^{(n)}\}_{n=1}^{N}$ is computed as
\begin{equation}
\text{\textsc{nmse}} = \frac{1}{N}\sum\limits_{n=1}^N\sqrt{\frac{\sum\limits_{d=1}^D\left(\widehat{\beta}_{d}^{(n)} - \beta_d^{(n)}\right)^2}{\sum\limits_{d=1}^D\left(\beta_{d}^{(n)}\right)^2}}
\end{equation}
In sparse coding it is usually more important to obtain the correct locations of spikes (i.e zeros) and slabs (i.e. non-zeros) in the estimator than to minimise \textsc{nmse}. The problem is therefore viewed as a skewed two-class classification problem where the number of spikes is higher than the number of slabs. F measure is used to evaluate the accuracy of such problems. 
F measure is computed as the harmonic mean of precision and recall
\begin{equation}
\text{F measure} = 2\dfrac{\text{precision}\cdot\text{recall}}{\text{precision} + \text{recall}}
\end{equation} 
where precision is the fraction of estimated slab locations that are correct, recall is the fraction of true slab locations among all predicted slab locations.
We demonstrate the performance on small datasets to highlight that the proposed algorithm can infer accurate predictions when the dataset size is not sufficient for \textsc{lista} to learn.

@@

In the Figure \ref{fig:number_of_layers_synthetic} prediction performance for different number of layers $L$ is presented. The observation size is set as $K=50$. In both experiments the performance of the Bayesian \textsc{lista} is better than the performance of \textsc{lista}. Although the baseline \textsc{ista} and \textsc{fista} show better performance in terms of F measure, Bayesian \textsc{lista} has lower \textsc{nmse}.

In the Figure \ref{fig:unsersampling_synthetic} prediction performance for different observation sizes $K$ is presented. The number of layers is set as $L=4$ which is the value of the poorest performance of Bayesian \textsc{lista} according to the previous experiment. Bayesian \textsc{lista} provides better results than {lista} for $K \ge 30$ and better results than other algorithms for $K \ge 70$.

@@

Figures \ref{fig:mnist_100} and \ref{fig:mnist_250} present quality on the validation set with dictionaries of size $100$ and $250$.  When $K=100$ Bayesian \textsc{lista} shows better performance that other algorithms and on $K=250$ it  converges to similar values as \textsc{lista}.

@@

%\begin{figure}[t]
%\includegraphics[width=\columnwidth]{loss}
%\caption{Validation \textsc{nmse}}
%\label{fig:validation}
%\end{figure}
\begin{figure}[t]
\subfloat[$\beta$ posterior mean]{\includegraphics[width=0.33\columnwidth]{posterior_mean}}~
\subfloat[$\beta$ posterior std]{\includegraphics[width=0.33\columnwidth]{posterior_std}}~
\subfloat[$\beta$ posterior spike indicator]{\includegraphics[width=0.33\columnwidth]{posterior_spike_indicator}}
\caption{Posterior for the digit 7.}
\label{fig:posterior_distribution}
\end{figure}
\begin{figure}[t]
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_0}}~
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_1}}~
\subfloat[]{\includegraphics[width=0.33\columnwidth]{posterior_sample_2}}
\caption{Samples from the posterior for the digit 7.}
\label{fig:posterior_samples}
\end{figure}

@@

The performance on the actively updated train data is compared with the performance on the randomly updated train data, when at every iteration the random element is moved from pool set to train set. Both algorithms of new data sampling were run with $20$ random seeds and their performance was averaged. Figure \ref{fig:active_learning_mnist} demonstrates their performance.The uncertainty sampling has some advantage compared to random sampling, this means that the posterior variance is adequately estimated.
\begin{figure}[t]
\subfloat[\textsc{nmse}]{\includegraphics[width=\columnwidth]{graphics/active_mnist/nmse_validation}} \\
\subfloat[F measure]{\includegraphics[width=\columnwidth]{graphics/active_mnist/f_measure_validation}}
\caption{Active learning on \textsc{mnist}.}
\label{fig:active_learning_mnist}
\end{figure}

\section{\uppercase{Discussion}}
\label{sec:discussion}
To the best of our knowledge, this is the first implementation of Bayesian deep sparse coding algorithm. Although there are works on Bayesian sparsity in context of neural networks \citep{he2017bayesian}, they are not the Bayesian neural networks in the same sense as Bayesian \textsc{lista} but rather the interpretation of the sparse Bayesian  learning algorithm as the long short-term memory network. We find not only correct predictions but also useful posterior estimates for the predictions that show how the model is confident in its decisions.
